---
title: "R Optimization HW1"
output: html_notebook
---

<br>
<br>
  
###Define decision variables:

***

**Regular Time** <br>
t_1 = # tables | f_1 = # floors | c_1 = # ceilings | p_1 = # pendants

**Overtime** <br>
t_2 = # tables | f_2 = # floors | c_2 = # ceilings | p_2 = # pendants

**Adertising** <br>
t_ad = proportion spent on table ads <br>
f_ad = proportion spent on floor ads <br>
c_ad = proportion spent on ceiling ads <br>
p_ad = proportion spent on pendant ads <br>

***

<br>
<br>

###Objective function: <br>

***

**Maximize:**

38t_1 + 49f_1 + 38c_1 + 68p_1 + 
         36t_2 + 47f_2 + 35c_2 + 65p_2 -
         10,000t_ad - 10,000f_ad - 10,000c_ad -10,000p_ad
         
***

<br>
<br>

###Subject To: <br>

***

**Min constraints:** <br>
t_1 >=0 | f_1 >= 0 | c_1 >= 0 | p_1 >= 0 <br>
t_2 >=0 | f_2 >= 0 | c_2 >= 0 | p_2 >= 0 <br> 
t_ad >= 0 | f_ad >= 0 | c_ad >= 0, p_ad >= 0 <br>

**Capacity:** <br>
t_1 + f_1 <= 100,000 | c_1 + p_1 <= 90,000 <br>
t_2 + f_2 <= 25,000 | c_2 + p_2 <= 24,000 <br>

**Potential Sales:**<br>
t_1 + t_2 - 7,200*t_ad <= 60,000 <br>
c_1 + c_2 - 2,000*c_ad <= 20,000 <br>
f_1 + f_2 - 8,000*f_ad <= 100,000 <br>
p_1 + p_2 - 5,250*p_ad <= 35,000 <br>

**Max Advertising Pct:** <br>
t_ad <= 1 | f_ad <= 1 | c_ad <= 1, p_ad <= 1 <br> 


**Ad Budget:** <br>
10,000t_ad + 10,000c_ad + 10,000f_ad + 10,000p_ad <= 18,000

***

###Set up the constraints and objective function to be solved 

```{r, warning = F, message = F}
library('lpSolve')

#variables:   t_1, t_2, f_1, f_2, c_1, c_2, p_1, p_2, t_ad,  f_ad,  c_ad,  p_ad
obj_coef <- c(38, 36, 49,  47, 38, 35, 68,  65, -10000, -10000, -10000, -10000)

basic_constraints <- rbind(diag(12),diag(12)[9:12,])

obj_names <- c('t_1', 't_2', 'f_1', 'f_2', 
               'c_1', 'c_2', 'p_1', 'p_2',
               't_ad', 'f_ad', 'c_ad', 'p_ad') 

colnames(basic_constraints) <- obj_names 

remain_constraints <- matrix(0,9,12)
remain_constraints[1,c(1,3)] <- 1
remain_constraints[2,c(5,7)] <- 1
remain_constraints[3,c(2,4)] <- 1
remain_constraints[4,c(6,8)] <- 1

remain_constraints[5,c(1,2,9)] <- c(1,1,-7200)
remain_constraints[6,c(3,4,10)] <- c(1,1,-2000)
remain_constraints[7,c(5,6,11)] <- c(1,1,-8000)
remain_constraints[8,c(7,8,12)] <- c(1,1,-5250)
remain_constraints[9,c(9,10,11,12)] <- c(10000,10000,10000,10000)

constraint_lhs <- rbind(basic_constraints,remain_constraints)
constraint_rhs <- c(0,0,0,0,0,0,0,0,0,0,0,0, # 12 greater than 0
                    1,1,1,1, # 16 less than 1 (max pct spent on ads)
                    100000, #17 less than 100,000 (Dept 1 regular capacity)
                    90000, #18 less than 90,000 (Dept 2 regular capacity)
                    25000, #19 less than 25,000 (Dept 1 overtime capacity)
                    24000, #20 less than 24,000 (Dept 2 overtime capacity)
                    60000, #21 less than 60,000 (Max table demand)
                    20000, #22 less than 20,000 (Max floor demand)
                    100000,#23 less than 100,000 (Max ceiling demand)
                    35000, #24 less than 35,000 (Max pendant demand)
                    18000) #25 less than $18,000 (Max ad budget)
                    
constraint_inequalities <- c('>=','>=','>=','>=','>=','>=','>=','>=','>=','>=','>=','>=',
                             '<=','<=','<=','<=',
                             '<=',
                             '<=',
                             '<=',
                             '<=',
                             '<=',
                             '<=',
                             '<=',
                             '<=',
                             '<=')
integers <- 1:8
```


```{r}
lp(direction = 'max', 
   objective.in = obj_coef, 
   const.mat = constraint_lhs,
   const.dir = constraint_inequalities,
   int.vec = integers,
   const.rhs = constraint_rhs,
   compute.sens = 1) -> lp_model

names(lp_model$solution) <- obj_names
```
<br>

###Questions:

***

**a) What is an optimal output plan for the company?**
```{r, echo = F}
lp_model$solution
```
<br>
**b) For each department, what is the marginal value of additional overtime capacity? (i.e. if we are able to increase the overtime capacity by one unit, how much impact would this have on theobjective function)**

```{r, echo=  F}
max_len <- length(lp_model$duals)- 12
duals <- lp_model$duals[1:max_len]
```

```{r, echo = F}
cat('An increase of one additional OT capacity in Department 1 results in an increase of ', duals[19], ' dollars')
```

```{r, echo = F}
cat('An increase of one additional OT capacity in Department 2 results in an increase of ', duals[20], ' dollars')
```

<br>
**c) What is the marginal value of additional advertising dollars?**
```{r, echo = F}
ad_dollars <- duals[25]

cat('For every additional advertising dollar spent we gain an additional ', ad_dollars, ' dollars')

```
<br>
**d) What is the marginal value of additional sales for each product?**

```{r, echo = F}
cat('The marginal value of an additional table =', duals[21],', floor =',duals[22], ', ceiling =', duals[23],', pendant =',duals[24])

```

***









