---
title: "Association Analysis Homework"
author: "Will Burton"
date: "October 8, 2016"
output: html_document
---

####Prompt <br>
The restaurant owner would like to add wine suggestions to the menu next to each meat and analyze her  
customer ordering habits in general. She is also interested in showing a picture of a popular meal of hers in  
an ad in Gourmet Diner magazine. Your job is to address her interests and also to report anything else of  
interest in terms of customer ordering habits, all in a nice summary report that is complete but to the point.  
 <br>
 <br>
 
####Load necessary packages and data
```{r, message = F, warning = F}
library(arules)
library(arulesViz)
library(dplyr)
library(tidyr)
setwd('C:/Users/Will/Documents/MSA/fall/data_mining/homework_1/data')
data<- read.csv('restaurantData.csv', stringsAsFactors = T)
```

####interest 1.   Find associations between wine and meat

```{r, warning = F, message = F}
#Since there was more than one meat, wine, and side per order number, the data needs to be split
#into seperate order numbers

#only look at meat and wine
data %>%
  dplyr::filter(type == 'Meat' | type == 'Wine') -> meat_wine 
  
#renumber to order numbers to have a single meat and wine combination per order number 
meat_wine %>% 
  mutate(orderNumber = sort(rep(seq(1,nrow(meat_wine)/2),2))) %>%
  mutate(orderNumber = factor(orderNumber)) %>% 
  tidyr::spread(type , order)-> meat_wine

#turn the resulting dataset into a dataset of class transaction
meat_wine <- as(meat_wine, 'transactions')

#Now we apply functions from the arules package to perform association analysis

#apriori algorithm
#target is association rules,
#minimum length of association is two factors
#min confidence = .01
#min support = .01
apriori_rules <- apriori(meat_wine, list(target = 'rules', supp=0.01, minlen = 2, conf = .01))  

#Return support for pairs:
head(inspect(sort(apriori_rules, by = 'lift')),10)

plot(apriori_rules, cex = 1)
```


####Interest 2.   Most popular meal 
<br>
```{r warning = F, message = F}
#To find the most popular meal, we concatenate side, wine, and meat and look at counts.
# The top 5 meals are returned
data %>% 
  mutate(orderNumber = sort(rep(seq(1,nrow(data)/3),3))) %>% 
  tidyr::spread(type,order) %>% 
  mutate(meal_combos  = paste(Meat, Side, Wine, sep = ', ')) %>% 
  group_by(meal_combos) %>% 
  summarise(Number_of_orders = n()) %>% 
  arrange(desc(Number_of_orders))-> meat_wine_side
  
head(data.frame(meat_wine_side))
```
The most popular meal combo is:
Pork Tenderloin, Roasted Root Veg, Cantina Pinot Bianco with 1154 orders,
82 orders higher than the next highest order
