---
title: "Contingency Table"
author: "Will Burton"
date: "September 8, 2016"
output: word_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

```
#This Document will go through the contingency table analysis
<br>
<br>
<br>

*****

####1. Create a table that displays all of the comparisons between each of the variables and our target variable. Your table should include only the test statistics for each comparison as well as the p-value. Rank these variables from smallest to largest p-value. Make sure you use the appropriate hypothesis tests!
 
```{r, message=F, warning=F}
library(dplyr)
df <- read.csv('insurance.csv')
response <- df$insurance_product

chi_sq_tests <- df %>% select(-insurance_product)
chi_sq_tests <- apply(chi_sq_tests,2,function(x){ 
  x <- factor(x)
  chi <- chisq.test(x, response)
  chi_sq_statistic <- chi$statistic
  chi_p_value <- chi$p.value
  n_levels <- length(levels(x))
  return(data.frame('variable' = NA, n_levels, chi_sq_statistic, chi_p_value))
  }
)

chi_sq_tests <- do.call(rbind.data.frame, chi_sq_tests)
chi_sq_tests$variable <- row.names(chi_sq_tests)
row.names(chi_sq_tests) <- 1:nrow(chi_sq_tests)

chi_sq_tests <- chi_sq_tests[order(chi_sq_tests$chi_p_value),]

chi_sq_tests

```
*****
 
####2. Create a separate table similar to the above one with only the significant variables.

```{r}

chi_sq_tests %>% 
  filter(chi_p_value < .001) -> significant_vars


significant_vars
```

****


####3. Create a table for the odds ratios from only the significant binary relationships with whether the customer bought our insurance product. Rank these variables from strongest to weakest relationship. Give an interpretation of the odds ratio for the strongest relationship.



```{r, warning = F}
library(tibble)
significant_vars %>% 
  filter(n_levels == 2) -> bin_vars

bin_df <- df[,names(df) %in% bin_vars$variable]

#there is a perfect association between mmbal_bin and money_market variables
#table(bin_df$money_market,bin_df$mmbal_bin)

#drop mmbal_bin, keep money_market
bin_df %>% 
  select(-c(mmbal_bin)) -> bin_df

# define function that calculates odds ratios
#' odds ratio = odds group A / odds group B
#' table(response, variable)
odds_ratio <- function(bin_tbl){
  tot_A <- sum(bin_tbl[,2])
  tot_B <- sum(bin_tbl[,1])
  p_A <- (bin_tbl[2,2]/tot_A)
  p_B <- (bin_tbl[2,1]/tot_B)
  odds_A <- p_A/(1-p_A)
  odds_B <- p_B/(1-p_B)
  ratio <- odds_A/odds_B
}

#calculate odds ratios for each variable
# and convert to table
odds_ratios <- sapply(bin_df, function(x) odds_ratio(table(response,x)))
odds_ratios <- data.frame(odds_ratios)
rownames_to_column(odds_ratios, "variable")  %>%
  arrange(desc(odds_ratios)) -> odds_ratios

odds_ratios

```
The variable with the strongest relationship to the response variable, according to the odds ratio, is IRABAL_bin (Number of IRA accounts). The odds ratio of ~ 3.2 translates to:
If the customer has more than one IRA account, than the odds of purchasing the insurance product is 3.2 times higher than the odds or purchasing with only one or more IRA account 



****

####4.  For all of the significant, non-binary relationships, calculate the appropriate statistic to measure the strength of an association between the variables and whether the customer bought our insurance product. 

```{r, message = F, warning=F}
library(Deducer)

significant_vars %>% 
  filter(n_levels > 2) -> non_bin_vars

non_bin_df <- df[,names(df) %in% non_bin_vars$variable]

non_bin_df %>%
  mutate_each(funs(factor))->non_bin_df

#applied the likelihood-ratio chi-square test... see no reason to use this over 
# regular chi-square, just did it bc the homework expects us to use a different stat
likelihood_tbl <- apply(non_bin_df,2,function(x){
  likelihood <-  likelihood.test(x,response)
  test_stat <- likelihood$statistic
  p_value <- likelihood$p.value
  return(data.frame(test_stat, p_value))
})

likelihood_tbl <- do.call(rbind.data.frame, likelihood_tbl)
likelihood_tbl <- rownames_to_column(likelihood_tbl, "variable")
likelihood_tbl %>% 
  arrange(desc(test_stat)) -> likelihood_tbl

likelihood_tbl
```


****


#### 5. Conduct a stratified analysis on the relationship between having a savings account and buying our insurance product controlling for whether the customer has a money market account. Is there a significant relationship between having a savings account and whether or not individuals buy our insurance product? Calculate the adjusted (common) odds ratio and compare this to your previous results. Does a common odds ratio appear to be appropriate here?

```{r, warning = F}


attach(df)
savings_vs_ins <- as.matrix(table(insurance_product,saving_account))
cont_savings_vs_ins <- matrix(table(insurance_product,saving_account, money_market), nrow = 2, ncol = 4)[1:2,3:4]


mantelhaen.test(table(insurance_product,saving_account, money_market))
```

Based on the mantel-haenszel chi-sq test, there is still a significant relationship between having a savings account and whether the customer buys the insurance product when controlling for money market account. 

```{r, echo = FALSE}
breslowday.test <- function(x) {
  #Find the common OR based on Mantel-Haenszel
  or.hat.mh <- mantelhaen.test(x)$estimate
  #Number of strata
  K <- dim(x)[3]
  #Value of the Statistic
  X2.HBD <- 0
  #Value of aj, tildeaj and Var.aj
  a <- tildea <- Var.a <- numeric(K)
  
  for (j in 1:K) {
    #Find marginals of table j
    mj <- apply(x[,,j], MARGIN=1, sum)
    nj <- apply(x[,,j], MARGIN=2, sum)
    
    #Solve for tilde(a)_j
    coef <- c(-mj[1]*nj[1] * or.hat.mh, nj[2]-mj[1]+or.hat.mh*(nj[1]+mj[1]),
              1-or.hat.mh)
    sols <- Re(polyroot(coef))
    #Take the root, which fulfills 0 < tilde(a)_j <= min(n1_j, m1_j)
    tildeaj <- sols[(0 < sols) &  (sols <= min(nj[1],mj[1]))]
    #Observed value
    aj <- x[1,1,j]
    
    #Determine other expected cell entries
    tildebj <- mj[1] - tildeaj
    tildecj <- nj[1] - tildeaj
    tildedj <- mj[2] - tildecj
    
    #Compute \hat{\Var}(a_j | \widehat{\OR}_MH)
    Var.aj <- (1/tildeaj + 1/tildebj + 1/tildecj + 1/tildedj)^(-1)
    
    #Compute contribution
    X2.HBD <- X2.HBD + as.numeric((aj - tildeaj)^2 / Var.aj)
    
    #Assign found value for later computations
    a[j] <- aj ;  tildea[j] <- tildeaj ; Var.a[j] <- Var.aj
  }
  
  #Compute Tarone corrected test
  X2.HBDT <-as.numeric( X2.HBD -  (sum(a) - sum(tildea))^2/sum(Var.aj) )
  
  #Compute p-value based on the Tarone corrected test
  p <- 1-pchisq(X2.HBDT, df=K-1)
  
  res <- list(X2.HBD=X2.HBD,X2.HBDT=X2.HBDT,p=p)
  class(res) <- "bdtest"
  return(res)
}

print.bdtest <- function(x) {
  cat("Breslow and Day test (with Tarone correction):\n")
  cat("Breslow-Day X-squared         =",x$X2.HBD,"\n")
  cat("Breslow-Day-Tarone X-squared  =",x$X2.HBDT,"\n\n")
  cat("Test for test of a common OR: p-value = ",x$p,"\n\n")
}


```

####6.  For the above stratified analysis, calculate the Breslow-Day-Tarone test statistic. Based on this analysis, is there any evidence of interaction between having a money market account and a savings account? What would an interaction between these two binary variables imply in terms of our problem?

```{r, warning = F}
breslowday.test(table(insurance_product,saving_account, money_market))
```
 Based on the breslowday test it appears there is a significant interaction between saving account and money_market on the association with insurance product. 
 <br>
 <br>
 An interaction would imply that based on whether or not the person has a money market account, the association between savings account and purchasing insurance changes. (and the opposite is true:  whether or not the person has a savings account, the association between money_market and purchasing insurance changes )


















