---
title: "Explore Insurance Dataset"
author: "Will Burton"
date: "September 8, 2016"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# This document will define variables used in the dataset, and go through a short exploratory analysis.

##Variable Breakdown
*****
###**Binary variables:**
####checking_account, saving_account  <br> 
####retirement_account, money_market  <br>
####insurance_product <br>
1 = person has attribute <br>
0 = person does not have attribute
<br>
<br>

###**Ordinal variables (more than 2 levels) :**
####ddabal_bin (checking account balance binned)<br>
ddabal = 0 ddabal_bin = 1<br>
ddabal > 0 & ddabal <=100 ddabal_bin = 2<br>
ddabal > 100 & ddabal <=300 ddabal_bin = 3<br>
ddabal > 300 & ddabal <=750 ddabal_bin = 4<br>
ddabal > 750 & ddabal <=1250 ddabal_bin = 5<br>
ddabal > 1250 & ddabal <=2250 ddabal_bin = 6<br>
ddabal > 2250 & ddabal <=6000 ddabal_bin = 7<br>
ddabal > 6000 ddabal_bin = 8<br>

####depamt_bin (amount deposited in checking account)
depamt = 2                      depamt_bin = 1<br>
depamt > 2 & depamt <=700       depamt_bin = 2<br>
depamt > 700 & depamt <=2200    depamt_bin = 3<br>
depamt > 2200 & depamt <=6500   depamt_bin = 4<br>
depamt > 6500                   depamt_bin = 5<br>

####checks_bin (Number of checks written)
checks = 0                 checks_bin = 1<br>
checks > 0 & checks <=2    checks_bin = 2<br>
checks > 2 & checks <=4    checks_bin = 3<br>
checks > 4                 checks_bin = 4<br>

####savbal_bin (balance in savings account)
savbal = 0                       savbal_bin = 1<br>
savbal > 0 & savbal <=50         savbal_bin = 2<br>
savbal > 50 & savbal <=250       savbal_bin = 3<br>
savbal > 250 & savbal <=1250     savbal_bin = 4<br>
savbal > 1250 & savbal <=3000    savbal_bin = 5<br>
savbal > 3000 & savbal <=8000    savbal_bin = 6<br>
savbal > 8000                    savbal_bin = 7<br>

####teller_bin (Number of teller banking transactions)
teller = 0                 teller_bin = 1<br>
teller > 0 & teller <=3    teller_bin = 2<br>
teller > 3                 teller_bin = 3<br>

####atmamt_bin (amount withdrawn from ATM)
atmamt < 5                       atmamt_bin = 1<br>
atmamt > 5 & atmamt <=3750       atmamt_bin = 2<br>
atmamt > 3750                    atmamt_bin = 3<br>

####cdbal_bin (amount in CD)
cdbal < 500                      cdbal_bin = 1<br>
cdbal > 500 & cdbal <=9200       cdbal_bin = 2<br>
cdbal > 9200                     cdbal_bin = 3<br>

####irabal_bin (Number of IRA accounts)
irabal < 1             irabal_bin = 1<br>
irabal > 1             irabal_bin = 2<br>

####mmbal_bin (Amount in money market)
mmbal < 1000 mmbal_bin = 1<br>
mmbal > 1000 mmbal_bin = 2<br>
<br>
<br>


###**Categorical variables:**
####acctage_bin (age of oldest account)
acctage = 20     acctage_bin = 1<br>
acctage > 20     acctage_bin = 2<br>
acctage = .      acctage_bin = 3<br>

####invbal_bin (Amount in Investments)
invbal < 1000                    invbal_bin = 1<br>
invbal > 1000 & invbal <=10000   invbal_bin = 2<br>
invbal > 10000                   invbal_bin = 3<br>
invbal = 10000                   invbal_bin = 4<br>

####phone_bin (Number of telephone banking transactions)
phone = 0                 phone_bin = 1<br>
phone > 0 & phone <=1     phone_bin = 2<br>
phone > 1                 phone_bin = 3<br>
phone = .                 phone_bin = 4<br>

####area_classification
####branch_of_bank


*****



The following section will display each variable in a barplot along with the associated table below the histogram


```{r, message = FALSE, warning = F}

library(sas7bdat)
library(dplyr)
library(ggplot2)
library(tidyr)
# df <- read.sas7bdat('C:/Users/Will/Documents/MSA/Homeworks/HW 1 Logistic/insurance_hw1.sas7bdat')
# labels <- unlist(sapply(attr(df, "column.info"), function(x) x$label))
# labels<- gsub(" ","_", labels)
# colnames(df)[1:7] <- labels
# colnames(df) <- tolower(colnames(df))
# write.csv(df, 'insurance.csv', row.names = F)

df <- read.csv('insurance.csv')
str(df)
summary(df)
df %>% 
  mutate_each(funs(factor)) -> df

for(colname in names(df)){
  print(ggplot(df) + geom_bar(aes(x = eval(eval(parse(text = paste("df$", colname, sep = "")))))) +
          xlab(colname) + ggtitle(paste("barplot of ", colname)))
  print(table(eval(parse(text=paste("df$", colname, sep = "")))))
}

```
