---
title: "Logistic Report"
author: "Will Burton"
date: "September 14, 2016"
output: html_document
---
```{r, warning = F, message=F}
library(dplyr)
library(tibble)
library(ggplot2)
library(caret)
library(pROC)

df <- read.csv('../data/insurance.csv')
  df %>% 
  select(-mmbal_bin) %>% 
  mutate_each(funs(factor)) -> model_df
attach(model_df)

# var_levels <- sapply(model_df, function(x) length(levels(x)))
# var_levels <- var_levels[var_levels == 2]
# interactions_tbl <- model_df[,names(model_df) %in% names(var_levels)]
# 
# significance_tbl <- NULL
# 
# for(i in 1:(ncol(interactions_tbl)-1)){
#   for(j in (i+1):ncol(interactions_tbl)){
#     var1 <- interactions_tbl[,i]
#     var1_name <- names(interactions_tbl)[i]
#     var2 <- interactions_tbl[,j]
#     var2_name <- names(interactions_tbl)[j]
#     interactions <-glm(insurance_product~ var1 + var2 + I(as.numeric(var1)*as.numeric(var2)-1), family = binomial)
#     interactions <- summary(interactions)
#     interactions <- interactions$coefficients
#     if(nrow(interactions) == 3){next}
#     rownames(interactions) <- c('intercept', var1_name, var2_name, paste0(var1_name, '*', var2_name))
#     interactions <- data.frame(interactions)[4,]
#     interactions <- rownames_to_column(interactions, 'variable')
#     significance_tbl <- rbind(significance_tbl,interactions)
#   }
# }
# 
# significance_tbl %>% 
#   filter(Pr...z.. < .001) -> significance_tbl
# 
# significance_tbl
```

1. Using all of the variables (including Branch of Bank) find the best model possible
on the training data set. Use information from your first assignment, as well as
ROC curves (do not go above 2-way interactions).
```{r }
# #stepwise formula based on AIC
# #define full model
# full_logistic_mod <- glm(insurance_product ~ . +
#                            I(as.numeric(checking_account)*as.numeric(retirement_account)) +
#                            I(as.numeric(checking_account)*as.numeric(money_market)) +
#                            I(as.numeric(checking_account)*as.numeric(irabal_bin)) +
#                            I(as.numeric(saving_account)*as.numeric(money_market)) +
#                            I(as.numeric(retirement_account)*as.numeric(money_market)), data = model_df, family = binomial)
# #define empty model
# nothing <- glm(insurance_product~ 1, family = binomial)
# 
# 
# #preform backwards, forwards, and stepwise selection, while optimizing AIC
# backwards <-step(full_logistic_mod, trace = 0) # Backwards selection is the default
# forwards <- step(nothing,
#                  scope=list(lower=formula(nothing),upper=formula(full_logistic_mod)), direction="forward", trace = 0)
# stepwise <- step(nothing, list(lower=formula(nothing),upper=formula(full_logistic_mod)),
#                  direction="both",trace=0)
# 

#Took output from forward, backward, and stepwise, and added terms to satisfy model heirarchy


#define interaction terms (could use I(var1 * var2) within the model but wanted to reduce amount of code)
model_df$ret_mm <- as.factor((as.numeric(retirement_account)-1) * (as.numeric(money_market)-1))
model_df$che_mm <-as.factor((as.numeric(checking_account)-1) * (as.numeric(money_market)-1))
model_df$che_ret <-as.factor((as.numeric(checking_account)-1) * (as.numeric(retirement_account)-1))
model_df$sav_mm <-as.factor((as.numeric(saving_account)-1) * (as.numeric(money_market)-1))


forwards_model <- glm(insurance_product ~  savbal_bin + ddabal_bin + cdbal_bin + 
                        branch_of_bank + checks_bin + atmamt_bin + teller_bin + 
                        checking_account + money_market  + 
                        saving_account + retirement_account + 
                        ret_mm + che_mm + che_ret + sav_mm, data = model_df,
                        family = binomial) 

backwards_model <- glm(insurance_product ~  checking_account + retirement_account + 
                         money_market + branch_of_bank + ddabal_bin + checks_bin + 
                         teller_bin + savbal_bin + atmamt_bin + cdbal_bin +
                         saving_account +  che_ret + sav_mm, data = model_df,
                         family = binomial) 


stepwise_model <- glm(insurance_product ~ savbal_bin + ddabal_bin + cdbal_bin + branch_of_bank + 
                        checks_bin + atmamt_bin + teller_bin + checking_account + money_market +
                        retirement_account + saving_account + ret_mm +  che_ret + 
                        sav_mm, data = model_df,
                         family = binomial)

```
<br>

1. Show me the ROC curve for your final model. (Make sure the axes and titles look
professional.)

```{r , warning = F }
#calculate roc curves for each model
calculate_ROC <- function(model, response, name, probs = NULL, i = NULL, test = TRUE){
  df <- NULL
  tp_rates <- NULL
  fp_rates <- NULL
  probs <- if(is.null(probs)){predict(model, type= 'response')} else{probs}
  AUC <- as.numeric(auc(response, probs))
    for(threshold in 0:1000){
      preds <- ifelse(probs > (threshold/1000), 1,0)
      confusion_matrix <- confusionMatrix(preds, response)$table
      POS <- confusion_matrix[2,2]
      NEG <- confusion_matrix[1,1]
      FALPOS <- confusion_matrix[2,1]
      FALNEG <- confusion_matrix[1,2]
      tp_rate <- POS / (POS + FALNEG)
      fp_rate <- FALPOS / (NEG + FALPOS)
      tn_rate <- NEG / (NEG + FALPOS)
      SPECIFICITY  <- tn_rate
      SENSIT <- tp_rate
      M1SPEC <- fp_rate
      df <- rbind(df, data.frame(name, AUC,'PROB' = threshold/1000,
                               POS, NEG, FALPOS, FALNEG, SENSIT,
                               M1SPEC, youden_index = (SENSIT + SPECIFICITY - 1), 
                               accuracy = (POS + NEG)/(POS + NEG + FALPOS + FALNEG))) 
                               
    }
  return(df)
}  


#ROC train set
Youden <- calculate_ROC(forwards_model, as.numeric(insurance_product)-1, 'forwards')


#plot roc curve for final model
ggplot(Youden, aes(x = M1SPEC, y = SENSIT)) + geom_line(size = 1) + geom_abline(slope = 1, intercept = 0) + xlab('False Positive Rate') + 
  ylab('True Positive Rate') + ggtitle("ROC Curve For Optimal Logistic Model") + 
    annotate("text", label = paste0("AUC = ", round(mean(Youden$AUC),4)), x = .75, y = .25, size = 5, colour = "Black")
```
<br>
What is the maximum Youden Index?


```{r }

Youden %>% 
  filter(youden_index == max(youden_index)) ->  max_youden

Youden %>% 
  filter(accuracy == max(accuracy)) %>% 
  select(PROB) ->  max_accuracy

print(max_youden %>% select(PROB, youden_index))
```
<br>
Based on the cut-off level that you defined in the previous part, create a classification table of your results. How many false positives and false negatives do you have?

```{r }

vals <- max_youden %>% select(NEG, FALNEG, FALPOS, POS)
tbl <- matrix(as.numeric(vals), nrow = 2, ncol = 2, byrow = T)
tbl<- as.table(tbl)
rownames(tbl)<-c(0,1)
colnames(tbl) <- c(0,1)
names(dimnames(tbl)) <- c("Prediction", "Reference")

tbl
vals %>% select(FALPOS, FALNEG) -> vals
names(vals) <- c('False Positives', 'False Negatives')
```
<br>
Build a classification table of your results using your validation data set. 
```{r }

library(sas7bdat)
valid <- read.sas7bdat('../data/insurance_hw2_valid.sas7bdat')
  
cols <- unlist(sapply(attr(valid,'column.info'), function(x) x$label))
cols <- as.character(sapply(cols, function(x) gsub(' ','_', x)))
names(valid)[1:length(cols)] <- cols  
names(valid) <- tolower(names(valid))
valid %>% 
  select(-mmbal_bin) %>% 
  mutate_each(funs(factor)) -> valid
  
detach(model_df)
attach(valid)
valid$ret_mm <- as.factor((as.numeric(retirement_account)-1) * (as.numeric(money_market)-1))
valid$che_mm <-as.factor((as.numeric(checking_account)-1) * (as.numeric(money_market)-1))
valid$che_ret <-as.factor((as.numeric(checking_account)-1) * (as.numeric(retirement_account)-1))
valid$sav_mm <-as.factor((as.numeric(saving_account)-1) * (as.numeric(money_market)-1))


probs <- predict(forwards_model, valid, type = 'response')
preds_youden <- ifelse(probs > max_youden$PROB,1,0)
preds_accuracy <- ifelse(probs > max_accuracy[1,],1,0)
tbl_youden <- confusionMatrix(preds_youden, insurance_product)$table
tbl_youden

tbl_accuracy <- confusionMatrix(preds_accuracy, insurance_product)$table
tbl_accuracy

```
<br>
How many false positives and false negatives do you have? 

```{r}
out <- c(tbl_youden[2,1], tbl_youden[1,2])
names(out) <- c("False Positives", "False Negatives")
out 

```
<br>
What is your misclassification error rate? 

```{r}
#using youdens
(tbl_youden[1,1] + tbl_youden[2,2]) /  length(probs)

#using threshold for max accuracy
(tbl_accuracy[1,1] + tbl_accuracy[2,2]) /  length(probs)

```



  

