---
title: "Simulation_hw1"
author: "Will Burton"
date: "November 12, 2016"
output: html_document
---

# Simulation Homework 1


Simulate possible future values of 2017 drilling costs.<br> 
* Currently, only previous information is available for 1960 - 2007.
* Since the industry has changed tremendously over those decades, only
the information from 1990 - 2006 will be useful for this analysis. 2007
was an outlier and should be ignored.
*Instead of looking at the distribution of actual costs, the Company's 
analysts recommend simulating possible annual changes in costs to get to 2017. <br>
<br>

They have calculated geometric changes in the data set already, but they are open to  other options if you explain why you chose them.
* Instead of focusing on costs for oil, gas, and dry wells individually, the
Company's analysts recommend to treat them all equally and assume an average
cost applies to them all. (HINT: You should have 48 observations. 
Geometric changes from 1991 - 2006.)
<br>
<br>
Previously the Price Analysis group has worked under the assumption that these 
geometric changes from one year to the next follow a Normal distribution.
Use QQ-plots or formal tests to see if you agree.
<br>
<br>

Build a kernel density estimate of the distribution of geometric changes,
using the 48 observations described above.
<br>
<br>
Simulate possible future values of 2017 drilling costs under 
both the assumption of Normality as well as under the kernel density 
estimate you created (HINT: Run two simulations). Make a recommendation 
for which one you feel the company should use.
 
***
***
***

Start off loading the R packages and the data
```{r, warning = F, message = F}
library(tidyr)
library(dplyr)
library(ggplot2)
library(ks)
setwd('C:/Users/Will/Documents/MSA/fall3/simulation/data')
cost <- read.csv('cost.csv')
```
<br>
Before we combine the costs into one column, lets explore the data and
determine if we really can combine them into one distribution

```{r }
#####################################################
################ Exploration ########################
#####################################################

#2007 is an outlier and should be ignored
cost <- cost[cost$Date >=1991 & cost$Date < 2007,]
cost$oil_return <- as.numeric(as.character(cost$oil_return))
cost$gas_return <- as.numeric(as.character(cost$gas_return))
cost$dry_return <- as.numeric(as.character(cost$dry_return))

#look at different costs
cost %>% 
  select(Date, oil_well_cost, gas_well_cost, dry_well_cost) %>% 
  gather(key = 'type', value = 'cost', oil_well_cost,
         gas_well_cost, dry_well_cost)-> drill_types_cost
```

<br>
Examine drilling costs over time
```{r}
ggplot(drill_types_cost, aes(x = Date, y = cost, color = type)) + geom_line() + 
  geom_point() + ggtitle('Cost of Drilling Over Time by Type')
ggplot(drill_types_cost, aes(factor(type), cost)) + geom_violin()
ggplot(drill_types_cost, aes(factor(type), cost)) + geom_boxplot()
```
<br>
Now lets look at the geometric changes over time 
```{r }
#performing an anova on the costs across the different types,
# there is no difference among groups and the associated cost

#look at geometric changes in cost over time
cost %>% 
  select(Date, oil_return, gas_return, dry_return) %>% 
  gather(key = 'return', value = 'cost', oil_return,
         gas_return, dry_return) %>% 
  mutate(cost = as.numeric(cost))-> drill_types_pct
ggplot(drill_types_pct, aes(x = Date, y = cost, color = return)) + geom_line() + 
  geom_point() + ggtitle('Cost of Drilling Over Time by Type')
  
ggplot(drill_types_pct, aes(factor(return), cost)) + geom_violin()
```
<br>
Now we perform an ANOVA to see if there are any differences between the different types of drilling: If there are differences then we should not combine them into the same group
```{r}
fit <- aov(cost ~ type, data = drill_types_cost)
summary(fit)

#anova assumptions:
#1. equal variance across groups
#2. errors are normally distributed
#3. observations are independent  

#1. 
#check for equal variance
bartlett.test(drill_types_pct$cost, factor(drill_types_pct$return))
car::leveneTest(drill_types_pct$cost, factor(drill_types_pct$return))
ggplot(drill_types_pct, aes(factor(return), cost)) + geom_boxplot() +
  geom_point(position = position_jitter(width = 0.2))
#Both the levene's test and bartletts test failed to reject the null

#2. 
# Check for normality
qqnorm(drill_types_pct$cost[drill_types_pct$return == 'oil_return'])
qqline(drill_types_pct$cost[drill_types_pct$return == 'oil_return'])
nortest::ad.test(drill_types_pct$cost[drill_types_pct$return == 'oil_return'])

qqnorm(drill_types_pct$cost[drill_types_pct$return == 'gas_return'])
qqline(drill_types_pct$cost[drill_types_pct$return == 'gas_return'])
nortest::ad.test(drill_types_pct$cost[drill_types_pct$return == 'gas_return'])

qqnorm(drill_types_pct$cost[drill_types_pct$return == 'dry_return'])
qqline(drill_types_pct$cost[drill_types_pct$return == 'dry_return'])
nortest::ad.test(drill_types_pct$cost[drill_types_pct$return == 'dry_return'])
# This is the only one that appears to be significantly different form normal
# I'd say for the most part this assumption checks out


#3.
#Check independence 
  independence_check <- lm(oil_return ~ lag(oil_return), data = cost)
  summary(independence_check)
#seems to check out, the lag is not significant and there is no visual pattern

  
# after checking the assumptions now we can perform an anova
# performing an anova across the different types to see if there
# are any differences
fit <- aov(cost ~ return, data = drill_types_pct)
summary(fit) 

#The global F-test fails to reject the null that there is no difference between
# the means of each group. 


#####################################################
################ End of Exploration #################
#####################################################

```


It turns out there is no difference between groups. The global F-test fails to reject the null that alll means are equal, and all the assumptions are confirmed. Because of this we can add them all into the same distribution

```{r}
# since there is no difference between groups, lets add them all 
# into the same distribution

#since one wasn't normal, lets try the anderson darling on all
# and also the qqplot
return_pct <- drill_types_pct$cost
```

Lets check for normality in the new combined data (though sums of normals should still be normal)
```{r}
qqnorm(return_pct)
qqline(return_pct)
nortest::ad.test(return_pct)
#The anderson darling fails to rejects the null that the distribution is normal
hist(return_pct)
```
<br>
Based on the qqplot and the anderson darling test, I would agree that 
the changes come from a normal population
<br>
<br>
Now we will simulate the future costs using a kernal density extimated distribution and the normal distribution

```{r}
#calculate optimal bandwidth
bw <- density(return_pct,na.rm = T, bw="SJ-ste")$bw


#estimate kernal density 
density_function <- kde(x = return_pct, h = bw)

#calculate average cost
drill_types_cost %>% 
  filter(Date == 2006) %>% 
  select(cost) -> average_cost 
average_cost <- mean(average_cost[,1])


#simulation based on estimated density function
simulate_return <- matrix(rkde(n = 11000, fhat= density_function), nrow = 1000)
simulate_return <- data.frame(simulate_return)
names(simulate_return) <- paste0('y_', 2007:2017)
simulate_return$final_return = apply(simulate_return,1,sum)
simulate_return$final <- average_cost + (exp(simulate_return$final) * average_cost)

par(mfrow = c(1,1))
hist(simulate_return$final)



#simulation based on a normal distribution
return_std_dev <- sd(return_pct)
return_mean <- mean(return_pct)
simulate_norm_return <- matrix(rnorm(11000,
                                      mean = return_mean, 
                                      sd = return_std_dev),nrow = 1000)
simulate_norm_return <- data.frame(simulate_norm_return)
names(simulate_norm_return) <- paste0('y_', 2007:2017)
simulate_norm_return$final_return = apply(simulate_norm_return,1,sum)
simulate_norm_return$final <- average_cost + (exp(simulate_norm_return$final) * average_cost)
```

Now that both simulations are completed, lets compare:
```{r}
#compare both distributions
par(mfrow = c(1,2))
hist(simulate_norm_return$final, xlab ='2017 cost', 
     main = 'Simulated using normal distribution',
     xlim = c(0,10000), breaks = 10)
hist(simulate_return$final, xlab = '2017 cost',
     main = 'Simulated using calcluated distribution', 
     xlim = c(0,10000), breaks = 12)

both_df <- data.frame('normal_final' = simulate_norm_return$final, 
                      'calculated_final' = simulate_return$final)

both_df <- gather(both_df)

#They are very similar
ggplot(both_df, aes(value)) + 
  geom_histogram(aes(fill = key), position ='identity', alpha =0.5)
```

I think the company should use the normal distribution since the observed
distribution is not different than the normal distribution
The normal distribution is interpretable and easily explained

